---
title: "Simulated Medical Trial Testing"
author: "Preeti Agrawal"
date: "2024-12-27"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(broom)


```

# **Introduction**

This project evaluates the effectiveness of three drug groups: Paracetamol, Paracetamol+Codeine, and Paracetamol+Caffeine, in treating post-traumatic acute pain. The analysis will test hypotheses about differences in pain reduction and patient satisfaction among the groups.

```{r load-data}
# loading the dataset
clinical_data <- read.csv("Simulated_Data_Clinical_Trial.csv")
head(clinical_data)

```

```{r summary-stats}
# Overview of the dataset
summary(clinical_data)

glimpse(clinical_data)

# Check for missing values
sum(is.na(clinical_data))

```

```{r group-distribution}
# Distribution of patients across groups
clinical_data %>% 
  ggplot(aes(x = Group)) +
  geom_bar(fill = "skyblue") +
  labs(title = "Distribution of Patients by Group", x = "Group", y = "Count") +
  theme_minimal()

```

### Exploring Pain Levels

```{r pain-levels}
# Boxplot for VAS scores at admission and discharge
clinical_data %>% 
  ggplot(aes(x = Group, y = VAS_Admission, fill = Group)) +
  geom_boxplot() +
  labs(title = "VAS Admission Scores by Group", x = "Group", y = "VAS Admission") +
  theme_minimal()

clinical_data %>% 
  ggplot(aes(x = Group, y = VAS_Discharge, fill = Group)) +
  geom_boxplot() +
  labs(title = "VAS Discharge Scores by Group", x = "Group", y = "VAS Discharge") +
  theme_minimal()

```

## Methodology

We aim to compare the effectiveness of the three drug groups in reducing pain.

-   **Testing Method**: ANOVA (Analysis of Variance) is used to evaluate if there are statistically significant differences in VAS scores among the groups.
    -   **Why ANOVA**: Since we are comparing more than two groups, ANOVA is appropriate to test mean differences.

## Analysis

### Checking Assumptions

```{r assumptions}
# Check for normality
shapiro.test(clinical_data$VAS_Discharge[clinical_data$Group == "Paracetamol"])
shapiro.test(clinical_data$VAS_Discharge[clinical_data$Group == "Paracetamol_Codeine"])
shapiro.test(clinical_data$VAS_Discharge[clinical_data$Group == "Paracetamol_Caffeine"])

# Check for homogeneity of variances
bartlett.test(VAS_Discharge ~ Group, data = clinical_data)

```

With the assumptions of normality and homogeneity of variances satisfied, an **ANOVA test** is appropriate for comparing the means of the VAS_Discharge variable across the three groups.

### ANOVA Test

```{r anova-test}
# Perform ANOVA
anova_results <- aov(VAS_Discharge ~ Group, data = clinical_data)
summary(anova_results)

# Tukey's HSD post-hoc test
TukeyHSD(anova_results)
```

### **Key Findings**

-   **Paracetamol+Codeine** is significantly more effective in reducing pain (lower VAS_Discharge) compared to **Paracetamol**.

-   There is no significant difference in pain relief between **Paracetamol+Caffeine** and the other two groups.

### **Recommendations**

1.  Consider **Paracetamol+Codeine** as a potentially more effective treatment for post-traumatic acute pain, particularly when compared to Paracetamol alone.

2.  Further studies with larger sample sizes could provide more conclusive evidence, especially for comparisons involving **Paracetamol+Caffeine**.

3.  Evaluate side effects and patient satisfaction alongside pain relief to make holistic treatment decisions.
